---
title: "Data Wrangling Exercise 2: Dealing with missing values"
output: html_notebook
---


Loading packages necessary for solving this exercise.


```{r message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
```


0: Load the data in RStudio
Save the data set as a CSV file called titanic_original.csv and load into RStudio in a data frame.  Also create a new data frame called 'titanic_clean'.

```{r message=FALSE}
titanic_original <- read_csv("titanic_original.csv")
titanic_clean <- titanic_original
```


1: Port of embarkation
The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r}
titanic_clean$embarked[is.na(titanic_clean$embarked)] <- "S"
```


2: Age
You’ll notice that a lot of the values in the Age column are missing. While there are many ways to fill these missing values, using the mean or median of the rest of the values is quite common in such cases.  
1. Calculate the mean of the Age column and use that value to populate the missing values
2. Think about other ways you could have populated the missing values in the age column. Why would you pick any of those over the mean (or not)?

```{r}
avg = mean(titanic_clean$age, na.rm = TRUE)
titanic_clean$age[is.na(titanic_clean$age)] <- avg
```


3: Lifeboat
You’re interested in looking at the distribution of passengers in different lifeboats, but as we know, many passengers did not make it to a boat :-( This means that there are a lot of missing values in the boat column. Fill these empty slots with a dummy value e.g. the string 'None' or 'NA'
```{r}
titanic_clean$boat[is.na(titanic_clean$boat)] <- "None"
```


4: Cabin
You notice that many passengers don’t have a cabin number associated with them.
*Does it make sense to fill missing cabin numbers with a value?
*What does a missing value here mean?
You have a hunch that the fact that the cabin number is missing might be a useful indicator of survival. Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r message=FALSE}
titanic_clean$has_cabin_number <- 0
titanic_clean$has_cabin_number[!is.na(titanic_clean$cabin)] <- 1
```

My guess is that a missing cabin means either another type of passage (i.e. steerage) or a stowaway.


5: Submit the project on Github
Include your code, the original data as a CSV file titanic_original.csv, and the cleaned up data as a CSV file called titanic_clean.csv.

```{r}
write.csv(titanic_clean, file = "titanic_clean.csv", row.names = FALSE)
```

If row.names is NOT set to FALSE then a new column is written to the CSV, creating a type of index column.  If this is left in when R reads the CSV in the additional column will need to be removed. 
